---
title: "Citeseq surface demultiplexing with aocseq"
output: rmarkdown::html_vignette
author: Noah Crooks and Mae Woods
vignette: >
  %\VignetteIndexEntry{vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width=6,fig.height=3
)
```

## Citeseq and cell surface labelling
In addition to loading and analyzing standard scRNA-seq data, aocseq is capable of handling data that is multiplexed using an arbitrary number of Hash Tagged Oligos (HTOs) and cell surface protein antibody oligos - see https://www.10xgenomics.com/analysis-guides/demultiplexing-and-analyzing-5%E2%80%99-immune-profiling-libraries-pooled-with-hashtags. These data are handled similarly to non-multiplexed data, but an additional demultiplexing step is added, such that cells are assigned to groups based on their hashtag counts and the probabilities generated by the GMM Demux algorithm. Demultiplexing allows single cells to be labelled by their samples and surface protein expression.

## Install Required Packages
Aocseq is designed for mixed cell type annotation of polyclonal populations of cells. Therefore, there is a running theme of genes of interest (gois) and cell types (clonotypes, receptor sequences, cell phenotypes). Gois are used and defined before analysis so that the data can be partitioned based on the transcript abundance of these genes. This example will show how a gene of interest can be included as metadata despite being removed by dimension reduction. Aocseq is compatible with the Seurat object data structure, and several other functions from the Seurat package can be imported and this example will show how cells that express a goi that is not used to create a 2D embedding can be easily displayed in any 2D plot, highlighting the ease of use to display low frequency cell subtypes. To begin load the required packages.

```{r}
library(aocseq)
library(ggplot2)
library(Seurat)
```


```{r}
suppressMessages(library(aocseq))
```

## Importing Data
Data import is handled by the CombineData function. For this vignette, we will use data from the paper "Chimeric Antigen Receptor-induced Antigen Loss Protects CD5.CART Cells From Fratricide Without Compromising On-target Cytotoxicity". These data contain 13278 T cells that have been co-cultured with CD5.CART cells from 3 healthy donors. The scRNA sequencing in this study was used to understand the transcriptional and cell phenotypic changes that occur in healthy T cells as a result of co-culture with CAR T cells that are directed toward a surface protein that is expressed on the surface of a T cell. The data can be downloaded from zenodo. Before running CombineData, download the example data and unzip the compressed files. 

```{bash, eval = F}
curl -O "https://zenodo.org/records/11286440/files/Figures-Rscripts.zip?download=1"

gunzip Figures-Rscripts.zip

cp Figures-Rscripts/Data/CD5/*.* Data
```

The raw data consists of cell hashing and gene expression transcripts and the cellranger processed FASTQ file output of feature barcode matrices can be imported. The number of variable are gex.path - this is the location of the feature barcode files, marker.gene - these are the gois, this example will show that CCL1 is expressed in donor 2 and 3 only and that although it is not included as a variable feature, it can still be displayed alongside dimension reduction. save.dir - this is the location to store a Seurat Object after demultiplexing. my.demultiplex.index - these are the surface values that you would like to run demultiplexing on, these hashtagged samples can be given names and in this example we call them my.names.hashtags. In this example, the first 7 slots in the antibody capture are surface protein and so we will not be demultiplexing these data. Finally, the number of hashtags used per sequencing run must be specified. This can be done with my.n.samples.ht.

```{r}
setwd("~/aocseqExample")
gex.path=c("C:/Users/mlwoods/Documents/aocseqExample/AOCSEQ_Example/SeqData/CD5")
marker.gene=c("IFNG","CRTAM","GZMB","CCR7","CCL1","SELL","CD28","CD27")
save.dir="C:/Users/mlwoods/Documents/aocseqExample/AOCSEQ_Example/RDS/CD5.rds"
my.demultiplex.index = c(8, 9, 10)
my.names.hashtags = c('donor_1', 'donor_2','donor_3')
my.n.samples.ht = 3
```
Using these inputs, CombineData will create a list of Seurat objects, one for each sample/hashtag combination. CombineData can perform QC filtering based on the nFeature_RNA_lower, nFeature_RNA_upper, and percent.mt_upper parameters. These parameters represent the lower and upper cutoffs for RNA feature counts, and the upper bound for mitochondrial RNA percentage, respectively. If set, these parameters produce figures with lines that show the cutoff and cells that are excluded. To speed up the example, we have set nvariable_features=500 but this can take any value and the default is 3000.
```{r}
CellData=CombineData(gex.path,
                     marker.gene,
                     file.saved=save.dir,
                     demultiplex=TRUE,
                     demultiplex.index=my.demultiplex.index,
                     nameshashtags=my.names.hashtags,
                     hashtags=3,
                     n.samples.ht = my.n.samples.ht,
                     nvariable_features=500,
                     QC_plots = TRUE)
```

Running CombineData will output the raw counts per sample and QC values. It will also provide a list of Seurat objects. In this case, we have three Seurat objects, one for each hashtag. As seen in the output from running the function CombineData, the goi CCL1 was not included in Seurat's sctransformed data and therefore it is not included in the variable features used for dimension reduction. Despite this, cells that express CCL1 can be subset in the data and displayed on a 2D UMAP embedding.

```{r}
my.ident.list=c("orig.ident","Threshold_CCL1","Threshold_SELL","Threshold_GZMB")
my.save.dir="AOCSEQ_Example"
```

```{r}
UMAPReduce(
  CellData[[2]],
  save.dir=my.save.dir,
  ident.list=my.ident.list
)
```

## Multimodal summary datasheets

Aocseq can be used to produce summary data sheets based on cell types of interest. In this data, cells have been labelled with surface antibodies and the data can be further demultiplexed to subset cell types of interest. This example will used the function Annotate cells to produce a summary data sheet that can then be used to visualize the differences in frequency of the cell types across the three donors with the function SegmentPlot.