---
title: "aocseq Vignette"
author: "Noah Crooks"
date: "2024-06-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
##Handling Hashtagged Data##
In addition to loading and analyzing standard scRNA-seq data, aocseq is capable of handling data that is multiplexed using an arbitrary number of HTOs. This data is handled similarly to non-multiplexed data, but an additional demultiplexing step is added, where cells are assigned to groups based on their hashtag counts and the probabilities generated by the GMM Demux algorithm.


###Install Required Packages###
Aocseq uses the Seurat object data structure, as well as several functions from the Seurat package. Seurat should be installed automatically along with aocseq, but we load it explicitly here to highlight that Seurat is central to the functionality of aocseq.

```{r}
library(aocseq)
library(Seurat)
```
###Importing Data###
Importing single-cell expression data and VDJ enrichment data is handled by the CombineData function. For this vignette, we will use data from the paper "Chimeric Antigen Receptor-induced Antigen Loss Protects CD5.CART Cells From Fratricide Without Compromising On-target Cytotoxicity" which is available from (website). These data contain 13278 T cells from 3 different patients. Before running CombineData, we need to define our parameters.
```{r}
gex.path=c("CD5")
vdj.path=c()
marker.gene=c("IFNG","CRTAM","TNF")  
threshold.cutoff=0.975
save.dir="CD5_2.rds"
index.control=1
n.samples=1
sample.name=c("CD5")
demultiplex.index = c(8, 9, 10)
nameshashtags = c('donor_1', 'donor_2','donor_3')
hashtags = 3
n.samples.ht = 1
n.hashtag.samples = 1
demultiplex=TRUE
c.index=1
sample.name=c("CD5")
nFeature_RNA_lower=100
nFeature_RNA_upper=10000
percent.mt_upper=5
verbose=TRUE
preset=1
```
Then we can run CombineData, which will create a list of Seurat object, one for each sample/hashtag combination. CombineData automatically performs QC filtering based on the nFeature_RNA_lower, nFeature_RNA_upper, and percent.mt_upper parameters.
```{r}
CellData=CombineData(gex.path,
  vdj.path=c(),
  marker.gene,
  threshold.cutoff,
  save.dir,
  index.control,
  n.samples,
  c.index,
  sample.name,
  preset,
  threshold.entry,
  demultiplex,
  demultiplex.index,
  nameshashtags,
  n.hashtag.samples,
  n.samples.ht,
  tenX_conversion="true",
  hashtags,
  nFeature_RNA_lower=100,
  nFeature_RNA_upper=10000,
  nvariable_features=3000,
  percent.mt_upper=5,
  verbose=TRUE)
```
After running CombineData, you should be left with a list of Seurat objects. In this case, we have three Seurat objects, one for each hashtag. We can then apply analysis and visualization to each object separately. While there is no VDJ data in these objects, we can use circos plots to visualize the different cell populations characterized by their surface marker expression.

