---
title: "Citeseq surface demultiplexing with aocseq"
output: rmarkdown::html_vignette
author: Noah Crooks
vignette: >
  %\VignetteIndexEntry{vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Citeseq and cell surface labelling
In addition to loading and analyzing standard scRNA-seq data, aocseq is capable of handling data that is multiplexed using an arbitrary number of Hash Tagged Oligos (HTOs) - insert hyperref. These data are handled similarly to non-multiplexed data, but an additional demultiplexing step is added, where cells are assigned to groups based on their hashtag counts and the probabilities generated by the GMM Demux algorithm. Demultiplexing allows single cells to be labelled by their samples and surface protein expression.

## Install Required Packages
Aocseq uses the Seurat object data structure, as well as several function from the Seurat package.

```{r}
library(aocseq)
library(ggplot2)
library(Seurat)
```


```{r}
suppressMessages(library(aocseq))
```

## Importing Data
Data import is handled by the CombineData function. For this vignette, we will use data from the paper "Chimeric Antigen Receptor-induced Antigen Loss Protects CD5.CART Cells From Fratricide Without Compromising On-target Cytotoxicity". These data contain 13278 T cells that have been co-cultured with CD5.CART cells from 3 healthy donors. The data can be downloaded from zenodo. Before running CombineData, dowload the example data and uncompress it. 

```{bash, eval = F}
curl -O "https://zenodo.org/records/11286440/files/Figures-Rscripts.zip?download=1"

gunzip Figures-Rscripts.zip

cp Figures-Rscripts/Data/CD5/*.* Data
```

We also need to define our parameters.

```{r}
gex.path=c("C:/Projects/SCsoftware/CD5")
marker.gene=c("IFNG","CRTAM","TNF")  
save.dir="CD5.rds"
my.sample.name=c("CD5")
my.demultiplex.index = c(8, 9, 10)
my.nameshashtags = c('donor_1', 'donor_2','donor_3')
my.n.samples.ht = 3
my.hashtags = 3
my.sample.name=c("CD5")
```
Then we can run CombineData, which will create a list of Seurat object, one for each sample/hashtag combination. CombineData automatically performs QC filtering based on the nFeature_RNA_lower, nFeature_RNA_upper, and percent.mt_upper parameters. These parameters represent the lower and upper cutoffs for RNA feature counts, and the upper bound for mitochondrial RNA percentage, respectively. To speed up the example, we have set nvariable_features=1000 but this can take any value and the default is 3000.
```{r, eval = F}
CellData=CombineData(gex.path,
                     marker.gene,
                     file.saved=save.dir,
                     sample.name=my.sample.name,
                     demultiplex=TRUE,
                     demultiplex.index=my.demultiplex.index,
                     nameshashtags=my.nameshashtags,
                     hashtags=my.hashtags,
                     n.samples.ht = my.n.samples.ht,
                     nvariable_features=1000,
                     QC_plots = TRUE)
```
After running CombineData, you should be left with a list of Seurat objects. In this case, we have three Seurat objects, one for each hashtag. While it is possible to run certain

