---
title: "Getting started with aocseq"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(aocseq)
```

## Trace sequencing

The package `"aocseq"` is a set of scripts that are designed to trace (among heterogenous cells) the expression of a set of genes that are functional bio-markers, in the sense that they are transiently co-expressed in a single cell and for a fixed duration of time.

## Loading the library and reading in the data with `CombineData`

Load the libraries for the package traceseq

```{r}
suppressMessages(library(aocseq))
```

The first step is to import single cell and (in the case of T cell activation and stimulation) VDJ contig files. In this example, 10X genomics feature barcode matrices are imported along with the contig_annotations_file.csv. Let's import a three hour stimulation assay. traceseq allows the selection of marker genes to compute their expression in a subset of cells

```{r, eval = F}
marker.gene=c("IFNG","TNF")
```
there are two conditions in our batch and the first condition is our control. We can specify a save directory for the RDS file.

```{r, eval = F}
save.dir="RDS/scAssay.rds"
```

and we have to specify the gene expression and VDJ path
```{r, eval = F}
gex.path=list("SeqData/gex/Control","SeqData/gex/CMV")
vdj.path=list("SeqData/vdj/Control/all_contig_annotations.csv","SeqData/vdj/CMV/all_contig_annotations)
```
              
After specifying the inputs we can run CombineData

```{r, eval = F}
scData <- CombineData(gex.path,vdj.path, marker.gene)
```

## Labelling subsets `AnnotateClonotypes`

AnnotateClonotypes summarizes and consolidates clonotypes (or cell types) across multiple samples. First we have to specify the Clonal Object

```{r, eval = F}
Clonal_Obs=readRDS("RDS/scData.rds")
```

We also need to specify and make a directory to store the summary data sheet

```{r, eval = F}
mypath="SummaryTables/scData"
mynames.spreadsheet = c("Control","CMV")
```

After this, we can call the function
```{r, eval = F}
AnnotateClonotypes(Clonal_Obs,goi="IFNG",goi.threshold=0.975,path=mypath,names.spreadsheet=mynames.spreadsheet)
```

## Classify cells based on their expression of the gene signature `ClassifyClonotypes`

Similarly to AnnotateClonotypes, you will need to specify a path to where you will save your data

```{r, eval = F}
myClonotypes=read_csv("SummaryTables/scDataIFNG/SummaryTable.csv")
mypath="SummaryTables/scData"
```

then run the function

```{r, eval = F}
ClassifyClonotypes(Clonal_Obs,myClonotypes,path=mypath)
```
